On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-02-24 MS.tex
Lizzy comments part 2

The entire dataset included studies differing in the growth measurement used (relative or absolute), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. However, these differences are clear criteria for assessing the quality of the data for addressing our hypotheses. Consequently, we adopted a two-stage approach where at first we were conservative by focusing exclusively on relative growth rate, across \underline{\textgreater} 10 species, under unstressed conditions, and where the source of the trait and growth is the same (expect for maximum height and seed mass, where a particular stage is by definition required to measure the trait). This dataset is called the ``ideal'' dataset. We then compared the results from the ``ideal'' dataset with those from the entire dataset - with both relative growth rate and absolute growth rates, and all other criteria relaxed.

@@

For SLA, the correlation between relative growth rate and the trait changed significantly with plant size in the manner expected (Fig. \ref{fig:fig3}a). The correlation between relative growth rate and SLA shifted from strongly positive at seedling stage ($\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero). The correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages. Model fit improved significantly when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3}a; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4}).

@@

We now report results from the entire dataset. Our aim is to determine if additional data alters or reinforces the above conclusions, by increasing replication. This entire dataset included correlations measured with both absolute growth rate and relative growth rate (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. traits could be measured on a different ontogenetic stage to growth data) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: Wood density-seedling, maximum height-sapling, maximum height-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

Results from the entire dataset reinforced our conclusions for both SLA and wood density, set apart alternative explanations for seed mass, and allowed us to compare correlations with growth along ontogenetic stages for Aarea. For SLA and wood density, the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ``ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across ontogenetic stage. The results were consistent with the prediction for a positive correlation with relative growth rate across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were still too few replicates to conclude undoubtedly that the correlation is absent for smaller sizes.

We had hypotheses for seed mass about a change in the correlation between trait and growth rate if absolute growth rate was used instead of relative growth rate. Effect size for seed mass differed significantly between relative growth rate and absolute growth rate at seedling stage; as expected the correlation with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI not overlapping). For wood density, some differences occurred according to the growth measurement used but they were not easily attributed to any single cause. In particular, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). Yet, the correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct}, and unlike others they used wood density values from adult plants together with growth from juveniles. So it was not possible to disentangle the effect of growth measurement from the effect of trait origin. For other traits, effect sizes were generally consistent between relative growth rate and absolute growth rate when both measurements were available.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assessed the coverage of the available data, to determine if biased sampling could have erroneously generated the above results.

@@

The meta-analysis reported here demonstrates that for several key traits, the influence of the trait on growth rate changes as plants grow from small to large. While the possibility has been raised recently that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}, this is the first time that specific mechanistic hypotheses have been corroborated against a wide array of existing studies. Meta-analysis allowed us to generalize and thus to reach a consensus about the effect of traits on growth, even if our conclusions were at times biased toward a particular growth stage, due to uneven coverage of the empirical literature across trait-growth-stage combinations. Growth rates were already known to change with size through ontogeny \citep{Condit:1993hd, Clark:1999ed, Herault:2011dd}, in other words individual plants follow a growth trajectory. Our results suggest species traits should be understood as influencing this trajectory and potentially re-ranking growth rates between individuals from different species as their size increases. This is a shift in outlook compared to much present thinking, whereby traits are seen as rendering species fast-growing or slow-growing throughout life.

@@

Meta-analysis showed a positive correlation between maximum height and relative growth rate for adult stage, but not for saplings and seedlings. The trend for tall stature species to grow faster at a given tree size than small stature species has been noted across several tree size classes and forest types (e.g. \citealt{Thomas:1996do,Poorter:2008iu,Wright:2010tp,Herault:2011dd,Ruger:2012jv,Iida:2014ep}). In \citet{Wright:2010tp}, maximum height was considered a weak predictor of growth rates of saplings; however it was the strongest predictor of growth rates of large trees. \citet{Poorter:2008iu} speculated that the maximum height-relative growth rate relationship among large trees might arise from including small tree species with maximum size not much larger than the 10cm dbh lower threshold used to define large trees. In our growth equations, maximum height affects growth via the fraction of energy diverted to reproduction. At a given tree size, species closer to their maximum height are expected to be allocating relatively less to vegetative growth. However this applies only at sizes where some of the species have begun to reproduce, and should not affect sapling and seedling growth rates. While the results of the meta-analysis were consistent with the expected pattern for maximum height, we noted that a limited number of replicates were available for both saplings (n=4) and seedling stages (n=1). Consequently, subsequent research is welcomed to validate this result.

@@

Many biotic and abiotic conditions such as light and water micro-environments can affect growth rates, and might modify correlations with traits in relation to plant size. For example, instead of a positive relationship that disappeared with plant size, \citet{Iida:2014ep} found SLA was negatively correlated with relative growth rate at small size and positively correlated with relative growth rate at larger sizes. Similarly, instead of a negative correlation between wood density and growth across all ontogenetic stages, some studies found a negative correlation at small size that disappeared when trees were larger \citep{Iida:2014ep,Iida:2014hq}. The authors explained the switch in directionality by changes in light and water condition up the vertical profile within a forest: larger trees reached a position in the canopy where they were released from light limitations, other resources such as water may have become limiting, and the performance of trees may then have been regulated by different traits. Note that in \citet{Iida:2014hq}, the shift in the correlation between growth and wood density occurred for trees larger than 22-26 cm dbh. This points to one of the possible limitations in our meta-analysis, where trees with diameter above 10 cm were pooled as ``adult''.

>>>>

The entire dataset included studies differing in the growth measurement used (relative or absolute), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. However, these differences are clear criteria for assessing the quality of the data for addressing our hypotheses. Consequently, we adopted a two-stage approach where at first we were conservative by focusing exclusively on relative growth rate, across strudies that included \underline{\textgreater} 10 species, under unstressed conditions, and where the source of the trait and growth is the same (expect for maximum height and seed mass, where a particular stage is by definition required to measure the trait). This dataset is called the ``ideal'' dataset. We then compared the results from the ``ideal'' dataset with those from the entire dataset - with both relative growth rate and absolute growth rates, and all other criteria relaxed.

@@

For SLA, the correlation between relative growth rate and the trait changed significantly with plant size in the manner expected (Fig. \ref{fig:fig3}a). The correlation between relative growth rate and SLA shifted from strongly positive at seedling stage ($\bar{r}=0.60 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero). The correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages. Model fit improved significantly when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3}a; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4}).

@@

We now report results from the entire dataset. Our aim is to determine if additional data alters or reinforces the above conclusions, by increasing replication. This entire dataset included correlations measured with both absolute growth rate and relative growth rate (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. traits could be measured on a different ontogenetic stage to growth data) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations that already had a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: Wood density-seedling, maximum height-sapling, maximum height-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

Results from the entire dataset reinforced our conclusions for both SLA and wood density and allowed us to compare correlations with growth across ontogenetic stages for Aarea and seed mass. For SLA and wood density, although the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ``ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across ontogenetic stage. The results were consistent with the prediction for a positive correlation with relative growth rate across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were still too few replicates to conclude undoubtedly that the correlation is absent for smaller sizes.

For seed mass, we had hypothesized that the correlation between trait and growth rate would change if absolute growth rate was used instead of relative growth rate. Effect size for seed mass differed significantly between relative growth rate and absolute growth rate at seedling stage; as expected the correlation with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI not overlapping). For wood density, some differences occurred according to the growth measurement used, but they were not easily attributed to any single cause. In particular, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). Yet, the correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct}, and unlike others they used wood density values from adult plants together with growth from juveniles. So it was not possible to disentangle the effect of growth measurement from the effect of trait origin. For other traits, effect sizes were generally consistent between relative growth rate and absolute growth rate when both measurements were available.

@@

Given our basic conclusion that reported correlations shifted with plant size in a manner consistent with expectations from theory, we assessed the coverage of the available data, to determine if biased sampling could have erroneously generated the above results.

@@

The meta-analysis reported here demonstrates that for several key traits, the influence of the trait on growth rate changes as plants grow from small to large. While the possibility has been raised recently that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}, this is the first time that specific mechanistic hypotheses have been corroborated against a wide array of existing studies. Meta-analysis allowed us to generalize and thus to reach a consensus about the effect of traits on growth, even if our conclusions were at times biased toward a particular growth stage, due to uneven coverage of the empirical literature across trait-growth-stage combinations. Growth rates were already known to change with size \citep{Condit:1993hd, Clark:1999ed, Herault:2011dd}, in other words individual plants follow a growth trajectory. Our results suggest species traits should be understood as influencing this trajectory and potentially re-ranking growth rates between individuals from different species as their size increases. This is a shift in outlook compared to much present thinking, whereby traits are seen as rendering species fast-growing or slow-growing throughout life.

@@

Meta-analysis showed a positive correlation between maximum height and relative growth rate for adult stage, but not for saplings and seedlings. The trend for tall stature species to grow faster at a given tree size than small stature species has been noted across several tree size classes and forest types (e.g. \citealt{Thomas:1996do,Poorter:2008iu,Wright:2010tp,Herault:2011dd,Ruger:2012jv,Iida:2014ep}). In \citet{Wright:2010tp}, maximum height was considered a weak predictor of growth rates of saplings; however it was the strongest predictor of growth rates of large trees. \citet{Poorter:2008iu} speculated that the maximum height-relative growth rate relationship among large trees might arise from including small tree species with maximum size not much larger than the 10cm dbh lower threshold used to define large trees. A possible mechanism  for Poorter et al's observation is captured by our growth equations: maximum height affects growth via the fraction of energy diverted to reproduction. At a given tree size, species closer to their maximum height are expected to be allocating relatively less to vegetative growth. However this applies only at sizes where some of the species have begun to reproduce, and should not affect sapling and seedling growth rates. While the results of the meta-analysis were consistent with the expected pattern for maximum height, we noted that a limited number of replicates were available for both saplings (n=4) and seedling stages (n=1). Consequently, subsequent research is welcomed to validate this result.

@@

Many biotic and abiotic conditions such as light and water micro-environments can affect growth rates, and might modify correlations with traits in relation to plant size. For example, for SLA, instead of a positive relationship with growth rates that disappeared with plant size, \citet{Iida:2014ep} found SLA was negatively correlated with relative growth rate at small size and positively correlated with relative growth rate at larger sizes. Similarly, instead of a negative correlation between wood density and growth across all ontogenetic stages, some studies found a negative correlation at small size that disappeared when trees were larger \citep{Iida:2014ep,Iida:2014hq}. The authors explained the switch in directionality by changes in light and water condition up the vertical profile within a forest: larger trees reached a position in the canopy where they were released from light limitations, other resources such as water may have become limiting, and the performance of trees may then have been regulated by different traits. Note that in \citet{Iida:2014hq}, the shift in the correlation between growth and wood density occurred for trees larger than 22-26 cm dbh. This points to one of the possible limitations in our meta-analysis, where trees with diameter above 10 cm were pooled as ``adult''.
